---
title: DATA 606 Data Project Proposal
author: JEYARAMAN RAMALINGAM
output:
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library("imputeTS")
library(bitops)
library(RCurl)

file_url <- getURL("https://raw.githubusercontent.com/jey1987/DATA606/master/Week6/2018-2010_export.csv")
india_trade <- data.frame(read.csv(text=file_url, header = TRUE),stringsAsFactors =FALSE)

india_trade_commodity_prep <- india_trade %>%
  select(Commodity,year) %>%
  filter(Commodity == "PHARMACEUTICAL PRODUCTS") %>%
  group_by(year,Commodity) %>%
  summarize(comm_cnt = n())

india_trade_export_prep <- india_trade %>%
  select(value,country,year) %>%
  group_by(year) %>%
  arrange(desc(value),.by_group=TRUE) %>%
  do(head(.,n=5))

  

```


### Research question 

**Below are two questions **

*1. How many number of exports India is performing over the years and how does the trend look like?*
*2. Which countries are top 5 contributors to India's export over time?*



### Cases 

**What are the cases, and how many are there?**
There are 137023 Observations and 5 variables


### Data collection 

**Describe the method of data collection.**

The data is collected by Kaggle and comprises of Indian export data from 2010 to 2018


### Type of study 

**What type of study is this (observational/experiment)?**

The Study is Observatory

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

Below is a link to the data source

https://www.kaggle.com/lakshyaag/india-trade-data


### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

Below are Response variable for two questions ,

1. Commodity Count (Numeric Variable)
2. Sum of export amount per Country in a given year


### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

1. HS2 column will act like independent variable and Year will act like an experimental variable

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r plot_sum , eval=TRUE}
summary(india_trade_commodity_prep)
dim(india_trade_commodity_prep)
head(india_trade_commodity_prep)

summary(india_trade_export_prep)
dim(india_trade_export_prep)
head(india_trade_export_prep)

ggplot(data=india_trade_commodity_prep, aes(x=year, y=comm_cnt, group=1)) +
  geom_line(linetype = "dashed",color="red")+
  geom_point() + theme_minimal() + theme(legend.position="top") +
  xlab("Year") + ylab("Number of Exports") + ggtitle("Number of Exports over years`") 

ggplot(data=india_trade_export_prep, aes(x=year, y=value, fill=country)) +
  geom_bar( stat="identity") + theme_minimal() + theme(axis.text.x = element_text(angle=90)) + 
  xlab("Countries") + ylab("Amount Exported") + ggtitle("Top 5 Amount Exported by Countries by Year ") 
```

